Residuals in zinbwave, 1000 most variable genes

### 2017-05-31 R/zinbwave_residuals.Rmd
residuals on 1000 and all genes

---
title: "Residuals in zinbwave"
author: "Fanny Perraudeau, Davide Risso"
date: "05/25/2017"
output: 
  html_document: 
    fig_height: 7
    fig_width: 7
    toc: yes
    code_folding: hide
    toc_float: yes
---

```{r options, echo=FALSE, results="hide",message=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
library(SummarizedExperiment)
library(scRNAseq)
library(zinbwave)
library(ggplot2)
library(magrittr)
library(matrixStats)
library(Rtsne)
library(scone)
library(RColorBrewer)
library(digest)
library(scone)
library(rARPACK)
set.seed(20)
```

@@

```{r datain}
load("../data/Expt4c2b_filtdata.Rda")
load("../data/E4c2b_slingshot_wsforkelly.RData")
```

@@

```{r}
plotBoxplot <- function(y, main = '', col, log = F){
  if (log) y = log2(y + 1)
  #median = apply(y, 1, median)
  #rle = apply(y, 2, function(x) x - median)
  boxplot(y, main=main, col=col, staplewex=0, outline=0, 
          border=col, xaxt='n')
  abline(h=0)
}

@@

compute.deviance.residuals <- function(Y, zinb){
  mu_hat = t(getMu(zinb)) 
  pi_hat = t(getPi(zinb)) 
  Y_hat = (1 - pi_hat) * mu_hat
  ll = compute.zinb.loglik(Y, zinb)
  sign = as.numeric(Y - Y_hat > 0)
  sign[sign == 0] = -1
  sign * sqrt(-2 * ll)
}

@@

# X intercept, no batch
Let's run zinbwave with K = 0 and X with an intercept only (no batch).
```{r}
print(system.time(zinb <- zinbFit(core, ncores = 3, K = 0)))
```

@@

@@

@@

```{r}
pca = prcomp(t(Rn))
par(mfrow = c(1,2))
plot(pca$x, col = batch, pch = 19,
     main="PCA of residuals\ncolor=batch")
plot(pca$x, col = clus.labels, pch = 19,
     main = "PCA of residuals\ncolor=cluster")
par(mfrow = c(1,1))
```
```{r}
Rn_order = Rn[, order(as.numeric(batch))]
col_order = as.numeric(batch)[order(as.numeric(batch))]
plotBoxplot(Rn_order, main = 'Boxplot of residuals\ncolor=batch',
            col = col_order)
```

@@

@@

# X batches
Let's run zinbwave with K = 0 and X with an intercept only (no batch).
```{r}
mod = model.matrix( ~ batch)
print(system.time(zinb_batch <- zinbFit(core, ncores = 3, K = 0,
                                  X = mod)))
```

## Naive residuals
```{r}
Rn = compute.naive.residuals(core, zinb_batch)
Rn[1:3,1:3]
```
```{r}
pca = prcomp(t(Rn))
par(mfrow = c(1,2))
plot(pca$x, col = batch, pch = 19,
     main="PCA of residuals\ncolor=batch")
plot(pca$x, col = clus.labels, pch = 19,
     main = "PCA of residuals\ncolor=cluster")
par(mfrow = c(1,1))
```

@@

@@

@@

>>>>

---
title: "Residuals in zinbwave, 1000 most variable genes"
author: "Fanny Perraudeau, Davide Risso"
date: "05/25/2017"
output: 
  html_document: 
    fig_height: 7
    fig_width: 7
    toc: yes
    code_folding: hide
    toc_float: yes
---

```{r options, echo=FALSE, results="hide",message=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
library(zinbwave)
library(matrixStats)
library(gplots) #heatmap.2()
library(clusterExperiment) #setBreaks()
set.seed(20)
NCORES = 8
```

@@

```{r datain}
load("../data/Expt4c2b_filtdata.Rda")
load("../data/E4c2b_slingshot_wsforkelly.RData")
de = read.csv('../data/oe_markers.txt', stringsAsFactors = F, header = F)
de = de$V1
```

@@

```{r}
plotBoxplot <- function(y, main = '', col, log = F){
  if (log) y = log2(y + 1)
  boxplot(y, main=main, col=col, staplewex=0, outline=0, 
          border=col, xaxt='n')
  abline(h=0)
}

@@

compute.deviance.residuals <- function(Y, zinb){
  mu_hat = t(getMu(zinb)) 
  pi_hat = t(getPi(zinb)) 
  Y_hat = (1 - pi_hat) * mu_hat
  ll = compute.zinb.loglik(Y, zinb)
  sign = 1*(Y - Y_hat > 0)
  sign[sign == 0] = -1
  sign * sqrt(-2 * ll)
}

@@

# X intercept, no batch
Let's run zinbwave with K = 0 and X with an intercept only (no batch).
```{r}
print(system.time(zinb <- zinbFit(core, ncores = NCORES, K = 0)))
```

@@

```{r}
de_1000 = de[de %in% rownames(Rn)]
length(de_1000)
head(de_1000)
```

@@

```{r}
breaks = setBreaks(Rn[de_1000, ], breaks = 0.99)
heatmap.2(Rn[de_1000, ], breaks = breaks,
          col = colorRampPalette(c('blue', 'yellow'))(51),
          scale = 'none', trace = 'none',
          ColSideColors = as.character(clus.labels),
          labCol = '', main = 'Naive Residuals')
```

@@

```{r}
heatmap.2(Rn[de_1000, ],
          col = colorRampPalette(c('blue', 'yellow'))(51),
          scale = 'none', trace = 'none',
          ColSideColors = as.character(clus.labels),
          labCol = '', main = 'Naive Residuals')
```

@@

```{r}
heatmap.2(Rp[de_1000, ],
          col = colorRampPalette(c('blue', 'yellow'))(51),
          scale = 'none', trace = 'none',
          ColSideColors = as.character(clus.labels),
          labCol = '', main = 'Pearson residuals')
```

@@

```{r}
heatmap.2(Rd[de_1000, ],
          col = colorRampPalette(c('blue', 'yellow'))(51),
          scale = 'none', trace = 'none',
          ColSideColors = as.character(clus.labels),
          labCol = '', main = 'Deviance residuals')
```

# X batches
Let's run zinbwave with K = 0 and X with an intercept only (no batch).
```{r}
mod = model.matrix( ~ batch)
print(system.time(zinb_batch <- zinbFit(core, ncores = NCORES, K = 0,
                                  X = mod)))
```

## Naive residuals
```{r}
Rn = compute.naive.residuals(core, zinb_batch)
Rn[1:3,1:3]
```

@@

```{r}
breaks = setBreaks(Rn[de_1000, ], breaks = 0.99)
heatmap.2(Rn[de_1000, ], breaks = breaks,
          col = colorRampPalette(c('blue', 'yellow'))(51),
          scale = 'none', trace = 'none',
          ColSideColors = as.character(clus.labels),
          labCol = '', main = 'Naive residuals')
```

@@

```{r}
heatmap.2(Rn[de_1000, ],
          col = colorRampPalette(c('blue', 'yellow'))(51),
          scale = 'none', trace = 'none',
          ColSideColors = as.character(clus.labels),
          labCol = '', main = 'Naive residuals')
```

@@

```{r}
heatmap.2(Rd[de_1000, ],
          col = colorRampPalette(c('blue', 'yellow'))(51),
          scale = 'none', trace = 'none',
          ColSideColors = as.character(clus.labels),
          labCol = '', main = 'Deviance residuals')
```
